MLPPI Wizard: An Automated Multi-level Partitioning Tool on Analytical Workloads

### 2017-03-03 mlppi-comsis17/paper.tex
MLPPI paper bibliography
Intro updated

%\section{Introduction}
%
%The text of the paper\dots

@@

Tables and materialized views in Teradata are hash distributed based 
on a user-specified column or set of columns called {\em primary index}. 
Each virtual processor, a unit of parallelism called {\em AMP} in the \hbox{Teradata} \hbox{relational} 
database management system (DBMS), 
receives a subset of the data and stores it in hash order.

Teradata users typically choose primary index fields, 
so that the data is evenly distributed among the AMPs. 
The primary index fields are also chosen to \hbox{reflect} join fields 
in workloads to accomplish cheaper local joins that do not require 
data shuffling.

@@

\begin{figure}[t]
\begin{center}
{\small
\begin{tabular}{l}
{\tt CREATE TABLE LINEORDER(} \\
\hspace{0.1in}{\tt LO\_ORDERKEY INTEGER,} \\
\hspace{0.1in}{\tt LO\_QUANTITY INTEGER,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT INTEGER} \\
{\tt )} \\
{\tt PRIMARY INDEX ( LO\_ORDERKEY )} \\
{\tt PARTITION BY (} \\
\hspace{0.1in}{\tt CASE\_N(} \\
\hspace{0.2in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.2in}{\tt NO CASE OR UNKNOWN),} \\
\hspace{0.1in}{\tt CASE\_N(}\\
\hspace{0.2in}{\tt LO\_QUANTITY < 25,} \\
\hspace{0.2in}{\tt LO\_QUANTITY >= 25 AND LO\_QUANTITY <= 30,} \\
\hspace{0.2in}{\tt LO\_QUANTITY > 30 AND LO\_QUANTITY <= 35,} \\
\hspace{0.2in}{\tt NO CASE OR UNKNOWN)} \\
{\tt );} \\
\end{tabular}
}
\end{center}
\vspace{-0.2in}
\caption{An Example of MLPPI Table\label{fig:exam}}
\end{figure}

@@

\begin{center}
\begin{tabular}{|c|p{6cm}|}\hline 
Partition & Condition \\ \hline
1		& {\scriptsize  {\tt LO\_DISCOUNT >= 7 \&\& LO\_QUANTITY < 25}} \\ \hline
2		& {\scriptsize  {\tt LO\_DISCOUNT >= 7 \&\& 25 <= LO\_QUANTITY <= 30}} \\ \hline
3		& {\scriptsize {\tt LO\_DISCOUNT >= 7 \&\& 30 < LO\_QUANTITY <= 35}} \\ \hline
4		& {\scriptsize {\tt LO\_DISCOUNT >= 7 \&\& LO\_QUANTITY no case}}\\ \hline
5       & {\scriptsize {\tt LO\_DISCOUNT no case \&\& LO\_QUANTITY < 25}} \\ \hline 
6		& {\scriptsize {\tt LO\_DISCOUNT no case \&\& 25 <= LO\_QUANTITY < 30}} \\\hline								  	        
7		& {\scriptsize {\tt LO\_DISCOUNT no case \&\& 30 < LO\_QUANTITY <= 35}} \\ \hline
8		& {\scriptsize  {\tt LO\_DISCOUNT no case \&\& LO\_QUANTITY no case}} \\ \hline
\end{tabular}
\end{center}

@@

\begin{center}
{\small
\begin{tabular}{rl}
$q$1:	& {\tt SELECT SUM(l.LO$\_$EXTENDEDPRICE*l.LO\_DISCOUNT)} \\ 
		& {\tt FROM LINEORDER l, DDATE d} \\
		& {\tt WHERE l.LO\_ORDERDATE = d.D\_DATEKEY} \\
		& {\tt AND d.D\_YEAR = `1993'} \\
		& {\tt AND l.LO\_DISCOUNT IN (1, 4, 5)} \\
        & {\tt AND l.LO\_QUANTITY <= 30} \\ 
		&						   \\	        
$q$2:	& {\tt SELECT c.C\_NATION, SUM(l.LO\_REVENUE)} \\ 
		& {\tt FROM CUSTOMER c, LINEORDER l} \\
		& {\tt WHERE l.LO\_CUSTKEY = c.LO\_CUSTKEY} \\
		& {\tt AND c.C\_REGION=`EUROPE'} \\
		& {\tt AND l.LO\_DISCOUNT >= 7} \\
		& {\tt AND l.LO\_QUANTITY >= 25 AND l.LO\_QUANTITY <= 35} \\
		& {\tt GROUP BY c.C\_NATION} \\
		& {\tt ORDER BY revenue desc}. \\
\end{tabular}
}
\end{center}

Let's now focus on the {\tt LINEORDER} fact table and 
the predicates involving {\tt LINEORDER} fields only. 
There are five constraints identified by the query number 
and the sequence number of the constraint in each query, as illustrated below.
%\vspace{-0.05in}	
\begin{center}
\begin{tabular}{|c|l|}\hline 
$q$1.1 & {\tt LO\_DISCOUNT IN (1,4,5)} \\ \hline
$q$1.2 & {\tt LO\_QUANTITY <= 30} \\ \hline
$q$2.1 & {\tt LO\_DISCOUNT >= 7} \\ \hline
$q$2.2 & {\tt LO\_QUANTITY >= 25}\\ \hline
$q$2.3 & {\tt LO\_QUANTITY <= 35} \\ \hline 
\end{tabular}
\end{center}
%\vspace{-0.05in}	
At this point the DBA needs to consider options 
based only on two fields and the five constraints. 
There are many possibilities from a {\em fine-grained} partition set 
to a loser definition.

Considering for the time being the column {\tt LO\_DISCOUNT}, 
one solution is to identify each value 
for the {\tt IN} predicate in $q$1.1 
and use $q$2.1 as is. This yields the following partitioning expression 
for {\tt LO\_DISCOUNT}:
\begin{center}
\begin{tabular}{l}
{\tt CASE\_N(} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 1,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 4,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 5,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.1in}{\tt NO CASE OR UNKNOWN} \\
{)}. \\
\end{tabular}
\end{center}
The above expression minimizes the size of the partitions 
by focusing exactly on the values required to satisfy the constraints, 
but creates a large number of small partitions.

@@

\begin{figure*}[htp!]
\centering
\includegraphics[scale=0.5]{architecture}
%\vspace{-0.2in}
\caption{The MLPPI Wizard Architecture\label{fig:arch}}
\end{figure*}

@@

The article is a substantial extension of prior work~\cite{Suh12}, 
and organized in the following way. 
Section~\ref{sec:algorithms} presents the multi-level \hbox{partitioning} 
algorithms for the wizard. 
In turn, Section~\ref{sec:analysis} conducts an analysis of the complexity of 
the algorithms. 
Section~\ref{sec:experiment} reports the performance evaluation results. 
Section~\ref{sec:related_work} states how \hbox{Teradata} MLPPI wizard 
is different from some existing tools proposed by other vendors. 
Lastly, Section~\ref{sec:conclusion} summarizes our discussion.

@@

In this section we describe in detail each phase of the wizard using the query set $Q$ provided in Section~\ref{sec:intro}.

@@

\bibliographystyle{splncs03}
\bibliography{paper}

>>>>

@@

Over the past decades much attention has been paid 
to improve the performance on analytical workloads in a data warehousing environment. 
Typically, a relational database management system (\hbox{DBMS}) 
has been exploited to process such analytical queries faster 
and assist its customers to make a timely business-decision in 
rapidly changing enterprise data warehousing.
Teradata DBMS has been a leading database product in this data warehousing marketplace 
that has been increasingly more competitive than ever.

To optimize the performance of analytical processing, 
tables and materialized views in the Teradata DBMS are hash-distributed based 
on a user-specified column or set of columns called {\em primary index}. 
Each virtual processor, a unit of parallelism called {\em AMP} in 
the \hbox{Teradata} \hbox{DBMS}, 
receives a subset of the data and stores it in hash order.

Users typically choose primary index fields in the DBMS, 
so that the data is evenly distributed among the AMPs. 
The primary index fields are also chosen to reflect join fields 
in workloads to accomplish cheaper local joins that do not require 
data shuffling.

@@

\begin{figure}[t]
\begin{center}
{\small
\begin{tabular}{l}
{\tt CREATE TABLE LINEORDER(} \\
\hspace{0.1in}{\tt LO\_ORDERKEY INTEGER,} \\
\hspace{0.1in}{\tt LO\_QUANTITY INTEGER,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT INTEGER} \\
{\tt )} \\
{\tt PRIMARY INDEX ( LO\_ORDERKEY )} \\
{\tt PARTITION BY (} \\
\hspace{0.1in}{\tt CASE\_N(} \\
\hspace{0.2in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.2in}{\tt NO CASE OR UNKNOWN),} \\
\hspace{0.1in}{\tt CASE\_N(}\\
\hspace{0.2in}{\tt LO\_QUANTITY < 25,} \\
\hspace{0.2in}{\tt LO\_QUANTITY >= 25 AND LO\_QUANTITY <= 30,} \\
\hspace{0.2in}{\tt LO\_QUANTITY > 30 AND LO\_QUANTITY <= 35,} \\
\hspace{0.2in}{\tt NO CASE OR UNKNOWN)} \\
{\tt );} \\
\end{tabular}
}
\end{center}
\vspace{-0.2in}
\caption{An Example of an MLPPI Table in the Teradata DBMS\label{fig:exam}}
\end{figure}

@@

\begin{center}
\begin{tabular}{|c|l|}\hline 
Partition & Condition \\ \hline
1		& {{\tt LO\_DISCOUNT >= 7 \&\& LO\_QUANTITY < 25}} \\ \hline
2		& {{\tt LO\_DISCOUNT >= 7 \&\& 25 <= LO\_QUANTITY <= 30}} \\ \hline
3		& {{\tt LO\_DISCOUNT >= 7 \&\& 30 < LO\_QUANTITY <= 35}} \\ \hline
4		& {{\tt LO\_DISCOUNT >= 7 \&\& LO\_QUANTITY no case}}\\ \hline
5       & {{\tt LO\_DISCOUNT no case \&\& LO\_QUANTITY < 25}} \\ \hline 
6		& {{\tt LO\_DISCOUNT no case \&\& 25 <= LO\_QUANTITY < 30}} \\\hline								  	        
7		& {{\tt LO\_DISCOUNT no case \&\& 30 < LO\_QUANTITY <= 35}} \\ \hline
8		& {{\tt LO\_DISCOUNT no case \&\& LO\_QUANTITY no case}} \\ \hline
\end{tabular}
\end{center}

@@

\begin{center}
{\small
\begin{tabular}{rl}
$q$1:	& {\tt SELECT SUM(l.LO$\_$EXTENDEDPRICE*l.LO\_DISCOUNT)} \\ 
		& {\tt FROM LINEORDER l, DDATE d} \\
		& {\tt WHERE l.LO\_ORDERDATE = d.D\_DATEKEY} \\
		& {\tt AND d.D\_YEAR = `1993'} \\
		& {\tt AND l.LO\_DISCOUNT IN (1, 4, 5)} \\
        & {\tt AND l.LO\_QUANTITY <= 30} \\ 
$q$2:	& {\tt SELECT c.C\_NATION, SUM(l.LO\_REVENUE)} \\ 
		& {\tt FROM CUSTOMER c, LINEORDER l} \\
		& {\tt WHERE l.LO\_CUSTKEY = c.LO\_CUSTKEY} \\
		& {\tt AND c.C\_REGION=`EUROPE'} \\
		& {\tt AND l.LO\_DISCOUNT >= 7} \\
		& {\tt AND l.LO\_QUANTITY >= 25 AND l.LO\_QUANTITY <= 35} \\
		& {\tt GROUP BY c.C\_NATION} \\
		& {\tt ORDER BY revenue desc} \\
\end{tabular}
}
\end{center}

Let's now focus on the {\tt LINEORDER} fact table and 
the predicates involving {\tt LINEORDER} fields only. 
There are five constraints identified by the query number 
and the sequence number of the constraint in each query, as illustrated below.
\vspace{-0.03in}	
\begin{center}
\begin{tabular}{|c|l|}\hline 
$q$1.1 & {\tt LO\_DISCOUNT IN (1,4,5)} \\ \hline
$q$1.2 & {\tt LO\_QUANTITY <= 30} \\ \hline
$q$2.1 & {\tt LO\_DISCOUNT >= 7} \\ \hline
$q$2.2 & {\tt LO\_QUANTITY >= 25}\\ \hline
$q$2.3 & {\tt LO\_QUANTITY <= 35} \\ \hline 
\end{tabular}
\end{center}
%\vspace{-0.05in}	
At this point the DBA needs to consider options 
based only on two fields and the five constraints. 
There are many possibilities from a {\em fine-grained} partition set 
to a loser definition.

Considering for the time being the column {\tt LO\_DISCOUNT}, 
one solution is to identify each value 
for the {\tt IN} predicate in $q$1.1 
and use $q$2.1 as is. This yields the following partitioning expression 
for {\tt LO\_DISCOUNT}:

\begin{center}
{\footnotesize
\begin{tabular}{l}
{\tt CASE\_N(} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 1,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 4,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 5,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.1in}{\tt NO CASE OR UNKNOWN} \\
{)}. \\
\end{tabular}
}
\end{center}
The above expression minimizes the size of the partitions 
by focusing exactly on the values required to satisfy the constraints, 
but creates a large number of small partitions.

@@

\begin{figure*}[htp!]
\centering
\includegraphics[scale=0.5]{architecture}
\vspace{-0.1in}
\caption{The MLPPI wizard architecture\label{fig:arch}}
\end{figure*}

@@

The article is a substantial extension of prior work~\cite{Suh12}. 
The rest of this article is \hbox{organized} in the following way. 
In the following section we propose a \hbox{multi-level} partitioning algorithm---consisting of a series of phases---used by our MLPPI wizard, 
which is applicable to any clustering or partitioning scheme in an arbitrary DBMS. 
In turn, we conduct a detailed analysis of the complexity of the algorithms. 
Next, we report the performance evaluation results. 
We then elaborate on how our wizard is distinguished from several existing tools proposed by other DBMS vendors. 
Finally, we conclude this article by summarizing our discussion.

@@

In this section we describe in detail each phase of the wizard using the query set $Q$ \hbox{provided} in Section~\ref{sec:intro}.

@@

{\footnotesize
\bibliographystyle{splncs03}
\bibliography{paper}
}
